---
title: "THREDDS data server"
author: "Hamze Dokoohaki"
date: "1/24/2022"
output:
  rmdformats::readthedown:
    self_contained: true
    thumbnails: false
    lightbox: true
    gallery: true
    highlight: haddock
    toc_depth : 4

---

This data server allows to read model inputs as well as observed data across borad regions for model initialization and model-data fusion purposes. This server can be combined with 'pSIMSCampaignManager' and 'pSIMSSiteMaker' R packages to create more robust simulations using initial condition, management and model parameters developed by our group or independent datasets.

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(leaflet)
library(leaflet.extras)
library(raster)
library(listviewer)
library(stars)
library(httr)
```

Necessary function for reading and subsetting NCDF files via OPeNDAP. 

```{r}

OpenDap_filereader <- function(url,variable, latVar, lonVar, ymax, ymin, xmin, xmax, datef=NULL) {
  # Finding all the lat/long before subsetting
  lon <- ncdf4::ncvar_get(ncdf4::nc_open(url),lonVar)
  lat <- ncdf4::ncvar_get(ncdf4::nc_open(url),latVar)
  
  
  #find the index of the bounds # estimate the difference between our points and lat/lon variable and
  # find indices that are closest to our x/y min/max
  ind.y.min <- which.min(abs(ymin - lat))
  ind.y.max <- which.min(abs(ymax - lat))
  ind.x.min <- which.min(abs(xmin - lon))
  ind.x.max <- which.min(abs(xmax - lon))
  #make the range of indices
  ydx <-  ind.y.min:ind.y.max
  xdx <- ind.x.min:ind.x.max
  
  #create the subset url
  subset.url <-
    paste0(
      url,
      '?',variable,'[',ind.y.min,':1:',ind.y.max,']',
      '[',ind.x.min,':1:', ind.x.max ,']'
    )
  
  print(subset.url)
 # browser()
  #read the subset url and replace lat/lon indices to real lat long
  p <- stars::read_ncdf(subset.url)
  p
}
```

Example of reading in maize planting date from the growing season file developed by AGMIP ; [This file can be found here in our TDS. ](http://192.17.59.89:8080/thredds/catalog/Allfiles/AGMIP_GROWING_SEASON.HARM.version1.25/catalog.html?dataset=testDatasetScan/AGMIP_GROWING_SEASON.HARM.version1.25/Maize_rf_growing_season_dates_v1.25.nc4) : 

## Reading file info using NCSS Dataset Description :

```{r,eval=FALSE}
XML::xmlToList("http://192.17.59.89:8080/thredds/ncss/Allfiles/AGMIP_GROWING_SEASON.HARM.version1.25/Maize_rf_growing_season_dates_v1.25.nc4/dataset.xml")
```

You can get access to the files on the server using few different services which here I quickly explain OPeNDAP and WMS. 

## OPeNDAP


```{r}

rast <- raster::raster("http://192.17.59.89:8080/thredds/dodsC/Allfiles/AGMIP_GROWING_SEASON.HARM.version1.25/Maize_rf_growing_season_dates_v1.25.nc4",varname='planting_day')

plot(rast)
```


Opendap with subset : 

The only downside with this method is that the `OpenDap_filereader` function calls the server and reads the original file 3 times which is not very efficient when hundreds/thousands of simulations attempt to call the server. A more efficient appraoch is presented under NCSS service. 

```{r message=FALSE, warning=FALSE}
url.f <- "http://192.17.59.89:8080/thredds/dodsC/Allfiles/AGMIP_GROWING_SEASON.HARM.version1.25/Maize_rf_growing_season_dates_v1.25.nc4"

# This function is defined on the top of this tutorial. 
subsetnc <- OpenDap_filereader(
  url = url.f,
  'planting_day',# Variable 
  'lat', # name of dim with lat values
  'lon',# name of dim with lon values
  ymax = 50, # defining a boundry 
  ymin = 25,
  xmin = -130,
  xmax = -70
)


plot(as(subsetnc, "Raster"))
```


## NetcdfSubset (NCSS ) 

We can also use NCSS to interactively  download a subset of the your NCDF. Another method for subsetting directly from the server (passing the task of subsseting files to the server) is POST a form on NCSS protocol as follows:

```{r, eval=FALSE}

u4 <-
  paste0(
    "http://192.17.59.89:8080/thredds/ncss/Allfiles/AGMIP_GROWING_SEASON.HARM.version1.25/Maize_rf_growing_season_dates_v1.25.nc4?var=planting_day&north=",
    50,
    "&west=",
    -130,
    "&east=",
    -70,
    "&south=",
    25,
    "&disableLLSubset=on&disableProjSubset=on&horizStride=1&accept=netcdf&addLatLon=true"
  )



resp <- POST(u4, body=list(), encode="form")

writeBin(content(resp),"test.nc4")

plot(raster::raster('test.nc4'))

```

## THREDDS R Package

```{r, eval=FALSE}
#devtools::install_github("bocinsky/thredds")

library(thredds)

cida <- "http://192.17.59.89:8080/thredds/catalog/Allfiles/catalog.html"

datasets <- tds_list_datasets(thredds_url = cida)
datasets
```

Choose your dataset (here is AGMIP_GROWING_SEASON.HARM.version1.25) and then list services for a specific file (here is Maize_ir_growing_season_dates_v1.25.nc4).

```{r, eval=FALSE}

datasets.plant <- tds_list_datasets(thredds_url = datasets[datasets$dataset =="AGMIP_GROWING_SEASON.HARM.version1.25/",]$path)

print(datasets.plant)

loca_services <- tds_list_services(datasets.plant[datasets.plant$dataset =="Maize_ir_growing_season_dates_v1.25.nc4",]$path)
loca_services

```

Choose your service (here I choose NCSS) and load the variables. 

```{r, eval=FALSE}
loca_ncss <- loca_services[loca_services$service == "NetcdfSubset",]$path

loca_vars <- tds_ncss_list_vars(ncss_url = loca_ncss)

loca_vars[,1:5]
```
```{r}
CCSM4 <- tds_ncss_download(ncss_url = loca_ncss,
                           out_file = paste0(tempdir(),"/plant.nc"),
                           bbox = sf::st_bbox(c(xmin = -130, xmax = -70, ymin = 25, ymax = 50)),
                           vars = loca_vars$name[4],
                           ncss_args = list(temporal = "all"))

plot(raster::raster(CCSM4))

```


## WMS

This service is extremely useful for quickly visualizing your nc files for simple comparison or visualizing and presenting model outputs.

```{r echo=TRUE}

leaflet() %>%
  addTiles %>% # Add default OpenStreetMap map tiles
  setView(lng = -90, lat = 40.0, zoom = 2) %>%
  addWMSTiles("http://192.17.59.89:8080/thredds/wms/Allfiles/AGMIP_GROWING_SEASON.HARM.version1.25/Maize_rf_growing_season_dates_v1.25.nc4",
    layers = "planting_day",
    options = WMSTileOptions(format = "image/png", transparent = TRUE),
    attribution = "Agornomic Forecasting Lab WMS server"
  ) %>%
  addWMSLegend(
    uri = "http://192.17.59.89:8080/thredds/wms/Allfiles/AGMIP_GROWING_SEASON.HARM.version1.25/Maize_rf_growing_season_dates_v1.25.nc4?REQUEST=GetLegendGraphic&LAYER=data_source_used&PALETTE=rainbow"
  )
```

