---
title: "Thredds data service"
author: "Hamze Dokoohaki"
date: "1/24/2022"
output:
  rmdformats::readthedown:
    self_contained: true
    thumbnails: false
    lightbox: true
    gallery: true
    highlight: pygments

---

This data server allows to read model inputs as well as observed data across borad regions for model initialization and model-data fusion purposes. 

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(leaflet)
library(leaflet.extras)
library(raster)
library(listviewer)
library(stars)
```

Necessary function for reading and subsetting NCDF files via OPeNDAP. 

```{r}

OpenDap_filereader <- function(url,variable, latVar, lonVar, ymax, ymin, xmin, xmax, datef=NULL) {

  # Finding all the lat/long before subsetting
  lon <- ncdf4::ncvar_get(ncdf4::nc_open(url),lonVar)
  lat <- ncdf4::ncvar_get(ncdf4::nc_open(url),latVar)
  
  
  #find the index of the bounds # estimate the difference between our points and lat/lon variable and
  # find indices that are closest to our x/y min/max
  ind.y.min <- which.min(abs(ymin - lat))
  ind.y.max <- which.min(abs(ymax - lat))
  ind.x.min <- which.min(abs(xmin - lon))
  ind.x.max <- which.min(abs(xmax - lon))
  #make the range of indices
  ydx <-  ind.y.min:ind.y.max
  xdx <- ind.x.min:ind.x.max
  
  #create the subset url
  subset.url <-
    paste0(
      url,
      '?',variable,'[',ind.y.min,':1:',ind.y.max,']',
      '[',ind.x.min,':1:', ind.x.max ,']'
    )
  
  print(subset.url)
 # browser()
  #read the subset url and replace lat/lon indices to real lat long
  p <- stars::read_ncdf(subset.url)
  p
}
```
## Reading file info using NCSS Dataset Description :

```{r}
 reactjson(XML::xmlToList("http://192.17.59.89:8080/thredds/ncss/Allfiles/AGMIP_GROWING_SEASON.HARM.version1.25/Maize_rf_growing_season_dates_v1.25.nc4/dataset.xml"))
```

## OPeNDAP

```{r}

rast <- raster::raster("http://192.17.59.89:8080/thredds/dodsC/Allfiles/AGMIP_GROWING_SEASON.HARM.version1.25/Maize_rf_growing_season_dates_v1.25.nc4",varname='planting_day')

plot(rast)
```

Opendap with subset : 

```{r}
url.f <- "http://192.17.59.89:8080/thredds/dodsC/Allfiles/AGMIP_GROWING_SEASON.HARM.version1.25/Maize_rf_growing_season_dates_v1.25.nc4"
subsetnc <- OpenDap_filereader(url=url.f, 'planting_day','lat','lon', ymax=50, ymin=25, xmin=-130, xmax=-70)


plot(as(subsetnc, "Raster"))
```


## NetcdfSubset (NCSS ) 

We can also use NCSS to interactivly to download a subset of the your NCDF

## WMS

This service is extremely useful for quickly visualizing your nc files for simple comparison or visualizing and presenting model outputs.

```{r echo=TRUE}

leaflet() %>%
  addTiles %>% # Add default OpenStreetMap map tiles
  setView(lng = -90, lat = 40.0, zoom = 2) %>%
  addWMSTiles("http://192.17.59.89:8080/thredds/wms/Allfiles/AGMIP_GROWING_SEASON.HARM.version1.25/Maize_rf_growing_season_dates_v1.25.nc4",
    layers = "planting_day",
    options = WMSTileOptions(format = "image/png", transparent = TRUE),
    attribution = "Agornomic Forecasting Lab WMS server"
  ) %>%
  addWMSLegend(
    uri = "http://192.17.59.89:8080/thredds/wms/Allfiles/AGMIP_GROWING_SEASON.HARM.version1.25/Maize_rf_growing_season_dates_v1.25.nc4?REQUEST=GetLegendGraphic&LAYER=data_source_used&PALETTE=rainbow"
  )
```