---
title: "Code Snippets"
author: "Hamze Dokoohaki"
date: "5/6/2022"
output:
  rmdformats::readthedown:
    self_contained: true
    thumbnails: false
    lightbox: true
    gallery: true
    highlight: haddock
    toc_depth : 4
---
```{r setup, include=FALSE}
options(warn=-1)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
message(.libPaths())
```

## Here we are collecting code snippets for a series of tasks, including simple visualiziation of maps etc. 

### Downloading Rotation maps for a county - Hamze 
```{r echo=TRUE, eval=FALSE}
library(lubridate)
library(ncdf4)
library(pSIMSSiteMaker)
library(pSIMCampaignManager)
library(furrr)
plan(multisession)
#-------
library(tidyverse)
library(raster)
library(stars)
library(maps)
library(sf)

county <- "lee"
years <- 2010:2020
state <- 'Illinois'

mapping_guid <- list(corn=1, soybeans = 0)
# ------------------------------------- Find the boundry of our county 
IL_counties <- st_as_sf( map('county', state, fill=TRUE, plot =FALSE)) %>%
  filter(grepl(county, ID)) %>%
  dplyr::select(-ID) %>%
  as(., "Spatial")
#------------------------------------ Generate empty raster layer and rasterize points for our county
county_raster <- raster(crs = crs(IL_counties), vals = 1,
                         resolution = c(0.01, 0.01), # 1 km by 1 km . every degree is ~ 100 km
                         ext = extent(IL_counties)) %>%
  rasterize(IL_counties, .)

#convert the raster to polygon to loop over each pixel
county_polys <- rasterToPolygons(county_raster)
#------------------------------------ For each pixel extract the rotation
rotations_df <- county_polys@polygons %>%
  future_map_dfr(possibly(function(x){
    
    xring <- x@Polygons[[1]]@coords[,1]
    yring <- x@Polygons[[1]]@coords[,2]
    
    pSIMSSiteMaker::historic_rotation_CropScape(xring, yring, years) %>%
      purrr::map_chr( ~ .x[[1]]) %>%
      as.data.frame() %>%
      mutate(years, 
             xcenter =  mean(xring),
             ycenter = mean(yring)
      ) %>%
      `colnames<-`(c("Crop", "Years", "xcenter", "ycenter"))
  }, otherwise = NULL), .progress = TRUE)


Rot_Rasters <- rotations_df %>%
  split(.$Years) %>%
  purrr::map(function(raster.oneyr){
    raster.oneyr$crp<- raster.oneyr$Crop %>%
      tolower() %>%
      purrr::map(~ ifelse(is.null(mapping_guid[[.x]]), -1, mapping_guid[[.x]]))
      
    raster::rasterFromXYZ(raster.oneyr[,c("xcenter","ycenter","crp")])
  }) %>%
  setNames(unique(rotations_df$Years))


names(Rot_Rasters) %>%
  purrr::map(~ plot(Rot_Rasters[[.x]], main=.x))

#saveRDS(Rot_Rasters, "Rot_Rasters.RDS")
#-------------------------------------------------------------------
MyVariable <- Rot_Rasters[[1]]
MyVariable[is.na(MyVariable)] <- -1

fname <-  "MyCampaign.nc4"
Create_Empty_Campaign(lat=unique(raster::xyFromCell(MyVariable, 1:length(MyVariable))[,2]),
                      lon=unique(raster::xyFromCell(MyVariable, 1:length(MyVariable))[,1]),
                      num_scen=50,
                      filename =fname)


new.values <- purrr::map(seq_along(num_scen),
                         ~Campaign_emptyMatrix("MyCampaign.nc4", raster::rasterToPoints(MyVariable)[,3])[[1]]
                         )

AddVar_Campaign("MyCampaign.nc4",
                Variable = list(Name=paste0('crid_',2),
                                Unit='Mapping',
                                missingValue=-99,
                                prec="float",
                                longname= "maize,soybean,fallow", 
                                value= new.values
                )
)


Inspect_Camp("MyCampaign.nc4")

########## Visualize 
library(tmap)

tm_shape(GetCamp_VarMatrix('MyCampaign.nc4','crid_2')$Raster[[1]])+
tm_raster("X1", palette = RColorBrewer::brewer.pal(3, "Set1"),
          title = "", alpha =0.85, style="cat")  + 
tm_layout(scale=1.2, bg.color = "white", legend.outside = TRUE)+
tm_grid()



```


### Reading and plotting yield for a county - Hamze 
```{r echo=TRUE, eval=FALSE}

library(rnassqs)
library(tidyverse)

api_key <- '' # https://quickstats.nass.usda.gov/api

nassqs_auth(key = api_key)
# Set parameters
params <- list(
  commodity_desc = c("CORN","SOYBEANS"),
  domaincat_desc = "NOT SPECIFIED",
  agg_level_desc = "COUNTY",
  state_alpha = "IL", # please make this an argument 
  year = 2010:2020 # please set this based on min max year in the simulated # you man need to read the simulated first
)

# Yields and Acres
yields <- nassqs_yields(params)%>%
  dplyr::select(commodity_desc, statisticcat_desc, unit_desc, county_name, year, Value) %>%
  filter(county_name == toupper("lee")) %>%
  mutate(value_kg_ha = Value * 62.77, 
         year=as.numeric(year)) %>%
  mutate(value_kg_ha_dry = purrr::map2_dbl(commodity_desc, value_kg_ha,  ~ ifelse(.x == 'CORN', .y * 0.845, .y * 0.87)))


yields %>%
  ggplot(aes(year, value_kg_ha_dry)) +
  geom_point(aes(color=commodity_desc), size=3)+
  geom_line(aes(color=commodity_desc))+
  facet_wrap(~commodity_desc, scales = "free")
```


### Reading APSIM xmls crop files and parsing the cultivars
```{r echo=TRUE, eval=TRUE}
library(dplyr)
library(xml2)



nodes <- read_xml('Maize.xml')
mmm <- xml_find_all(nodes, ".//Model")

mm <- xml_children(mmm)%>%
  xml2::xml_path()

# find cultivar tags and their paths
cultivars.path <- mm%>% 
   purrr::map(function(tmm){
    xml2::xml_children(mmm) %>%
      xml2::xml_find_all(tmm) %>%
      xml2::xml_attr('cultivar')
  }) %>%
  setNames(mm) %>%
  purrr::discard(~is.na(.))


names(cultivars.path) %>%
   purrr::map_dfr(function(pp){
    
    
    data.frame(
      tt_emerg_to_endjuv = xml2::xml_children(mmm) %>%
      xml2::xml_find_all(pp) %>%
      xml2::xml_children() %>%
      xml2::xml_find_all(paste0(pp,"/tt_emerg_to_endjuv ")) %>%
      xml2::xml_text() %>%
      as.numeric(), 
      tt_flower_to_maturity = xml2::xml_children(mmm) %>%
        xml2::xml_find_all(pp) %>%
        xml2::xml_children() %>%
        xml2::xml_find_all(paste0(pp,"/tt_flower_to_maturity ")) %>%
        xml2::xml_text() %>%
        as.numeric(),
      tt_flower_to_start_grain = xml2::xml_children(mmm) %>%
        xml2::xml_find_all(pp) %>%
        xml2::xml_children() %>%
        xml2::xml_find_all(paste0(pp,"/tt_flower_to_start_grain ")) %>%
        xml2::xml_text() %>%
        as.numeric()
    ) %>%
      mutate(Cultivar= basename(pp))
   
  }) %>%
  summary

```


### Creating pSIMS complient Met files form ERA5:
```{r echo=TRUE, eval=FALSE}
library(pSIMSMetMaker)
library(tidyverse)
setwd("C:/Users/hamzed/Downloads")
sites.path <- system.file("", 'pSIMS_extent.RData', package = "pSIMSMetMaker")
load(sites.path)
#------------------------------------------------- ERA - NEEDS LOGIN
#Login setup
#Create a free CDS user account by self registering.
#Once your user account has been verified you can get your personal user ID and key by visiting the user profile.
#This information is required to be able to retrieve data via the ecmwfr package
#https://github.com/bluegreen-labs/ecmwfr


  i <- 1
  user <- "10027"
  wf_set_key(user = user, key = "e5d0082f-ae4d-44d6-a1e0-3b581282033f", 'cds')
  
  # Extracting the tile(s) lat/lon info
  sites <- pSIMS.extent %>%
    dplyr::filter(name %in% c('0025/0046'))
  # Request ERA5 data
  era5_data <- era5_tile_request(user, "2017-01-01", "2021-12-31",
                                 sites$xmin[i], sites$xmax[i], sites$ymin[i], sites$ymax[i],
                                 fname=paste0(gsub("/", "_", sites$name[i]),".nc"))
  # Reformat ERA5 data
  era5_data_refmt <- reformat_ERA_data(era5_data)
  #Aggregate the ERA5 data
  era5_data_Agg <- ERA5_aggregate(era5_data_refmt)
  
  pSIMSMetMaker::write_met(era5_data_Agg[[1]], paste0('clim_', gsub("/", "_", sites$name[i]), '.tile.nc4'))
  


```

### Estimating ensemble weights post-hoc after assimilation:
```{r echo=TRUE, eval=FALSE} 
library(mvtnorm)
library(furrr)
plan(multisession)

Ens_Weighting <- function(SDA.out){

  SDA <- SDA.out %>% discard(~ length(.x)==1)
  
  Allw <- data.frame(Date = SDA %>% map_chr(~ .x$Obs$Date %>% as.character) %>% as.Date()) %>%
    mutate(id=1:n()) %>%
    mutate(Year=lubridate::year(Date)) %>%
    split(.$Year) %>%
    future_map_dfr(function(one.year){

      SDA.year <- SDA[one.year$id]
      # Estimate the wights for each year
      SDA.year %>%
        map_dfr(function(ss){
          flux.weights <- dmvnorm(ss$Forecast$X,
                                  mean =ss$mu.a,
                                  sigma = ss$Pa,
                                  log = FALSE)
          
          
      data.frame(Weights= scales::rescale(flux.weights), 
                 Date=ss$Obs$Date) %>%
        mutate(Ens= 1:n())
        })
    }, .progress = TRUE)
 
  Annual.W <- Allw %>%
    mutate(Year = lubridate::year(Date)) %>%
    group_by(Ens, Year) %>%
    summarise(
      Weight= sum(Weights)
    )
  
  Annual.W %>%
    split(.$Year) %>%
    map_dfr(function(yr){
     yr$ScaledW <- yr$Weight/sum(yr$Weight)
     yr %>%
       dplyr::select(-Weight)
    })
}

ss <- Ens_Weighting(SDA.out)

```

### Re-running broken simulations after pSIMS generates the simulations:

Find the directories with files "core*" . This file is usually created when a simulation breaks

```{bash , eval=FALSE}
#!/bin/bash

for i in $(find /scratch/users/tsrai/illinois_gallatin_SDA/illinois_gallatin_SDA/* -name "core*" ); do 

	DIRNAME=$(dirname $i)
	RUNPATH="${DIRNAME}/APSIM_Run.sh" 
	echo ${RUNPATH}
	cp -fr "APSIM_Run.sh" ${DIRNAME}
	qsub -v simpath=${DIRNAME} ${RUNPATH}
	echo "----------------------------------"
done


```

APSIM_Run.sh :

```{bash , eval=FALSE}

#!/bin/sh
#SBATCH --time=12:00:00  
#SBATCH --partition=aces
#SBATCH --mem=4000
#SBATCH --nodes=1
#SBATCH --ntasks-per-node=1
source ~/.bashrc
simpath=${simpath}

echo ${simpath}

singularity exec -B /projects/aces/tsrai/psims/Data:/pysims/data /projects/aces/hamzed/psims/Data/SingularityImg/pSIMSLatest.simg mono /Apsim79/Model/Apsim.exe "${simpath}/Generic.apsim"
```