---
title: "pSIMS Tutorial"
author: "Hamze Dokoohaki"
date: "4/5/2021"
output:
  prettydoc::html_pretty:
    theme: cosmo
    highlight: vignette
    toc: true
    toc_depth: 4
    

---


## Download Files used in this Tutorial

```{r echo=FALSE}
downloadthis::download_file("Maize_template.xml", "Maize_template", "Maize cultivar parameters", button_type = "primary")
downloadthis::download_file("exp_template.json", "exp_template", "Campgain json file", button_type = "primary")
```

***


```{r setup, include=FALSE}
options(warn=-1)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
```


This tutorial is only for the cases in which the user intends to ensemble crop parameters other than the crop cultivars. In that case, make sure that you `pysims` is updated to the latest version. You can find the latest version here :  [Link to the repository](https://github.com/AgronomicForecastingLab).

The key to use parameters other than cultivar for creating ensembles is to make sure your campaign json file includes the following tag:

```{json}
  "cultivarParams":{
	"minsw" : "0.00002"
  }
```

This tag is named `cultivarParams` and under this tag, you may include any parameter that needs to vary, although the real process of creating ensembles happen when creating the campaign ncdf using `pSIMCampaignManager`. 


The second step is to include the following tag inside your crop.xml (such as Maize.xml) for the parameter of interest:

```{xml}
$cultivarParams.minsw
```
The format of the above code follows the structure of your json name convention. In this case we are varying the `minsw` parameter inside the Maize.xml and the overall xml tag will look like below:

```{xml}
<minsw description="lowest acceptable value for ll">$cultivarParams.minsw</minsw>
```


The final last change is in the apsim template. In your apsim template the crop component should look like this :

```{xml}
      <$simulation.cropName>
            <ini>
               <filename>${simulation.cropName_f}.xml</filename>
            </ini>
      </$simulation.cropName>
```

The only different between this tag and a typical apsim templates is that the ensemble specific crop cultivar is `simulation.cropName_f` instead of `simulation.cropName`.


The rest will be taken care of by `pysims` by creating a separate Maize.xml for each simulation (Ensemble) and assigning them to their corresponding Maize  module. 