---
title: "pSIMS Tutorial"
author: "Hamze Dokoohaki"
date: "4/5/2021"
output:
  rmdformats::readthedown:
    self_contained: true
    thumbnails: false
    lightbox: true
    gallery: true
    highlight: haddock
    toc_depth : 4
    

---


## Download Files used in this Tutorial

```{r echo=FALSE}
downloadthis::download_file("LAI_Obs.csv", "LAI_Obs", "Sample LAI Measurments", button_type = "primary")
```

***

In this tutorial, I will be using `pSIMSSiteMaker` and `pSIMCampaignManager` packages to create an emulator similar to the  [Link to Emulator Tutorial](http://192.17.59.89:3838/pSIMSEmulator/). The only different is the use of Object Oriented Programming (OOP) implemented in the `pSIMSSiteMaker` package with the goal of encapsulating the routine data processing and GP development. This method is extremely useful for multi-site and specially hierarchical model calibration. This class developed `Gaussian Process` (GP) models as an emulator at every time point where there is observation. The `Simulator` object then would allow you generate a time series of GP model estimates for a given set of parameters. 

```{r setup, include=FALSE}
options(warn=-1)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
message(.libPaths())
```



```{r}
library(pSIMSSiteMaker)
library(tidyverse)
```

### Creating the emulator using the GP class

The first step is design your parameter space like below. Make sure you use the same name convention used below:

```{r}
params <- list (
  pars = c("tt_flower_to_maturity",
           "tt_emerg_to_endjuv",
           "tt_flower_to_start_grain"),
  Upper.limit = c(900, 300, 250),
  lower.limit = c(700, 200, 100),
  unit=c('C/day','C/day','C/day')
)

print(params)
```


**It is very important for the observed data to have a column named Date. This is particularly the same for the model outputs as well.** You have to also make sure that there is only one observation for each time period.

```{r}
readr::read_csv("LAI_Obs.csv")
```

You can easily create you emulator by creating an object of class GP, by providing 4 arguments to the `new` function from this class. Then the object created will include all the data, model output and a series of functions for running GPs over all the dates and a numerical optimization method. You can treat this object similar to how `lists` are used in `R` in addition to a series of functions that are able to use their internal data and model output.

```{r}

siteGP <- GP$new(DataPath="LAI_Obs.csv",
                 ModelOutPath="/home/hamzed/pSIMS/Emulator/Results",
                 Params=params,
                 BuildLikGP=FALSE,
                 GPType="GAM",
                 lenght.out.grid=4)

```

You can check you grid over parameter space using the `Grid` property.

```{r}

head(siteGP$Grid)

```

You can also use the `inspect_obs()` function within your object to inspect the observed data. A similar function is also developed for inspecting the model outputs.

```{r}
siteGP$inspect_obs()


```


```{r}
siteGP$inspect_ModelOut()


```

We can also double check the out files that are loaded into our GP object.

```{r}
head(siteGP$Model.out.files)

```

```{r}
names(siteGP$GPs)

```

`Simulator` is an important function exposed to the users where you can create simulations of `GP`estimates given a series of parameters within your parameter space. 

```{r eval=FALSE, include=TRUE}

siteGP$Simulator(newParam = data.frame(tt_flower_to_maturity=751,
                                       tt_emerg_to_endjuv=234,
                                       tt_flower_to_start_grain= 181))
```

Below we are visualizing model estimates (Using all the GPs) for a given set of parameter against the observed data.


```{r}
tmp_obs <- siteGP$obsData %>%
  `colnames<-`(c("Date", "Obs"))

siteGP$Simulator(newParam = data.frame(tt_flower_to_maturity=751,
                                       tt_emerg_to_endjuv=234,
                                       tt_flower_to_start_grain= 181))


```

`objfun` exposes the objective function used for numerical optimization for testing purposes. 

```{r}
siteGP$objfun(c(751.3503,234.3428, 181.5699),  tmp_obs)
```

Finally user can perform numerical optimization using the Shuffle Complex Evolution method simply by providing a series of initial conditions. 

```{r eval=FALSE, include=TRUE}
siteGP$Optim_sceuares(c(751.3503, 234.3428, 181.5699))
```



### Debugging GP class:

R6 generator objects have a method called debug() which will enable debugging for a method. This will affect all instances of the class that are created after the debug() is called.


```{r eval=FALSE, include=TRUE}
GP$debug("objfun")
```

This will be then used like below :

```{r eval=FALSE, include=TRUE}
s <- GP$new()
s$objfun()
# [Debugging prompt]


# Disable debugging for future instances:
GP$undebug("objfun")
```


To enable debugging for a method in a single instance of an object, use the debug() function (not the debug() method in the generator object).

```{r eval=FALSE, include=TRUE}
s <- GP$new()
debug(s$objfun)
s$objfun()
# [Debugging prompt]
#Use undebug() to disable debugging on an object’s method.

undebug(s$objfun)
```


