---
title: "Downloading VI in R from GEE"
author: "Hamze Dokoohaki"
date: "5/3/2021"
output:
  prettydoc::html_pretty:
    theme: cosmo
    highlight: vignette
    toc: true
    toc_depth: 4
    

---

In this tutorial, I will be using `L8S2` package to download vegetation indices in R. This tutorial only works on our private server, mainly because it leverages the docker image that is hosted on our server. Before I get into how to use this package, you can find this package here:

```{r}
#https://github.com/AgronomicForecastingLab/L8S2
```


```{r setup, include=FALSE}
options(warn=-1)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
```



```{r}
library(L8S2)
library(tidyverse)
```
```{r}
?DownloadL8S2
```
### Defining the parameters

```{r}
mysites <- data.frame(x= c(-88.20,-88.20),  #long
                      y= c(40.06,42.06),    #lat
                      ID= c('EnergyFarm',"RandomFarm") #Site ID
)

RS <- DownloadL8S2(mysites,  '2020-01-01', '2021-01-01', Indices = c("NDVI", "EVI", "SAVI", "OSAVI"))
```

Now we can inspect and visualize the outputs
```{r}
head(RS)

```


```{r}
RS %>%
  ggplot(aes(date, Value))+
  geom_point(aes(color=Satellite))+
  geom_line(aes(color=Satellite))+
  facet_wrap(Index~ ID)+
  theme(legend.position="top")

```


Of course you can use the `write.csv` function also to write this data.frame to the CSV format. 
